Learning Augmented Joint-Space Task-Oriented Dynamical Systems: \\ A Linear Parameter Varying and Synergetic Control Approach

### 2017-10-27 paper/main_RA-L_v1_nadia.tex
Created another _nadia.tex file with my changes.. Still have to finish Section 5.. discuss results.

\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{url}
\usepackage[font={footnotesize}]{caption}
\usepackage[font={footnotesize}]{subcaption}
\usepackage[usenames, dvipsnames]{color}
\usepackage{amsmath}
\renewcommand{\labelenumii}{\Roman{enumii}} 
\usepackage{authblk}
\usepackage[hidelinks]{hyperref}
% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
\usepackage{amssymb}

@@

%Lastly, we wish to capture joint-space behaviors in our dynamical system.
We make two decisions by the choice of behavior error metric \eqref{eq:optim_error}. First, the metric implicitly suggests that behaviors are defined by expert motions of the behavior, and that fulfilling a behavior means moving similarly to the experts' motion. Second, the type of feature of the trajectory that is being minimized is important. If instead the error feature were joint position distance $\sum_i\left \| q_{d,i} - q_i \right \|$ where $q_d$ is a vector of demonstrated positions, then executing a "behavior" would imply mimicking position, but not velocity. Alternatively, if the error feature were joint velocity direction $\sum_i \left | \frac{\dot{q}_{d,i}}{\left \| \dot{q}_{d,i} \right \|} - \frac{\dot{q}_i}{\left \| \dot{q}_i \right \|} \right |$ mimicking a "behavior" would involve following the motion profile, but not the motion's speed (so for example a slap and a push would exhibit the same behavior). By choosing the combined direction and magnitude of the joint velocity as the error, we are choosing to mimic the magnitude and direction of the motion, which \cite{gielniak2010stylized} suggests is visually most similar to the human definition of "joint motion style".
\section{Augmented \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem} \label{Sec:DS}
\label{sec:proposed_system}
\textcolor{red}{Change semi-definite to definite everywhere\\ Introduce 'behavior synergies' here\\}
We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the two criteria presented in \ref{eq:Lyp} and \ref{eq:optim_error}.
\begin{equation}
\label{eq:ds}
\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
\end{equation}
where $P\in \mathbb{R}^{d\times d}$, and $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $\mathcal{A}(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy. 
\begin{equation}
\label{eq:A_def}
\mathcal{A}(q) = \sum_{k=1}^{K}\theta_k(q)A_k 
\end{equation}
where $A_k\in \mathbb{R}^{m\times m} $ are the different synergies and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are the scheduling parameters\footnote{The scheduling parameters can be a function of time $t$, the states of the
system $q $ or external signals $d(t)$, i.e. $\theta_k (t, q(t), d(t))$. In this paper, we only consider it as a function of the states of the system. It is noteworthy that the presented stability proof can easily be extended for $\theta_k (t, q(t), d(t))$.} modulating each local synergy through time as well as space. Based on \eqref{eq:ds}, one can utilize different synergies in different regions, and compose them to create a more complex multi-behavior motion.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{Pic/A_comparison_positions.pdf}
\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $\mathcal{A}(q)$ (emphasizing the hip, knee, and ankle respectively), A: $\mathcal{A}(q)=diag(5,1,1) $, B: $\mathcal{A}(q)=diag(1,5,1)$, and C: $\mathcal{A}(q)=diag(1,1,5)$. On the left, the task-space traces of each motion. On the right, the time-scaled joint positions of each joint. Each motion tends to use its ``primary" joint most and uses the other available joints to compensate for what the primary joint cannot do.}
\label{fig:A_example}
\vspace{-15pt}
\end{figure}

Before proving that the proposed dynamical system satisfies all of the criteria, let us establish an intuitive understanding of the components of the system. While it may seem daunting at first, each of the elements has a straightforward explanation. One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t. the task target (warped according to P), and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}). The positive semi-definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig. \ref{fig:A_example} illustrates the effects of  $\mathcal{A}(q)$ on the generated motion. Thus the controller can be thought of as a proportional controller in joint space. Lastly, we refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(because it augments the task error, and thus the direction of motion) and $\mathcal{A}(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (because it augments the outputted joint velocities).
%When using each of these matrices, the controller will augment the joint velocities to favor using one joint more than the others (in $A_1$ the hip, in $A_2$ the knee, and in $A_3$ the ankle). This ''augmentation" is simply a matrix multiplication ($A\dot{q}_{raw}$). The components of the outputted velocity are linear combinations of the ''raw" un-augmented joint velocities, where the entries in $A$ define the precise ratios. We can see the joint positions resulting from our experiment over time on the right side of Fig. \ref{fig:A_example}. Each motion tends to use its ``primary" joint most (the joint where the velocity is heavily scaled up) and uses the other available joints to compensate for what the primary joint cannot do. For example, Motion 1 starts out using mostly the hip, but once moving the hip no longer brings the system closer to the goal, the system uses the other joints to fold inwards and reach the goal. Ultimately, all three controllers converge on their target, but each has a different path through joint space depending on which joints we choose to emphasize in the behavior matrix $A$. The off-diagonal terms of $A$ have a more nuanced contribution; e.g. $A_{1,2}$ defines the amount that elbow velocity should be converted into hip velocity.

@@

%By learning a position-dependent joint augmentation matrix $\mathcal{A}(q)$, we are able to execute the joint behaviors discussed earlier by augment the relative velocities of our joints. \eqref{eq:A_def} constructs our overall $\mathcal{A}(q)$ as a linear combination of $A_k~\forall k\in\{1,\dots,K\}$ (each weighted by $\theta_k(q)$), where one can enforce different local joint behaviors in different regions, and compose them to create a more complex multi-behavior motion. In this paper the joint behavior matrix $\mathcal{A}(q)$ is determined using only the joint position $q$, though one could potentially use other information such as joint velocity $\dot{q}$, environment force or target task position $x^*$.\\

\section{Learning \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem in Synergy Space} 
\label{Sec:Learning}

The behavior of the JT-DS algorithm can be best understood through the lens of synergy control \cite{petrivc2013muscle}. In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the actuation-space that are sufficient to accurately recreate the desired robotic behaviors. In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and $\mathcal{A}(q)$ represents the resulting motion constructed from a superposition of different synergies (through \eqref{eq:ds}).

@@

\subsection{Estimating the Synergy Matrices}
Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $\mathcal{A}(q)$ as a linear combination of local $A_k$ matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows: 
\begin{equation}
\label{eq:A}
\mathcal{A}(q) = \frac{\sum_{k=1}^K A_k \pi_k\mathcal{N}(\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}.
\end{equation} 
Notice the resemblance of \eqref{eq:A} to the Nadaraya-Watson kernel estimator \cite{nadaraya1964regress,watson1964regress}\footnote{The Nadaraya-Watson kernel estimator is used to estimate an unknown regressive function $m(x) = \mathbb{E}\{Y|X\}$, which takes the general form of $\widehat{m}(x) = \frac{\sum_{i=1}^n y_i \mathcal{K}(x,x_i)}{\sum_{i=1}^n \mathcal{K}(x,x_i)}$ where $\mathcal{K}(x,x_i)$ is a kernel function denoting the distance or similarity of $x_i$ to the given location $x$. \cite{nadaraya1964regress,watson1964regress}} with a Gaussian pdf as its kernel function. Hence \eqref{eq:A} can be considered a type of kernel estimator, with the key distinction that the weighting functions $\theta_k'(\phi(q))$ are not determined by individual points (as in the original Nadaraya-Watson kernel estimator) but by the components of a GMM, similar to the weighting functions derived in Gaussian Mixture Regression (GMR) \cite{sung2004gmr}.

@@

\newpage
\section{Experimental Validation} \label{Sec:Exp}

\subsection{JT-DS Learning  Evaluation}

The learning algorithm presented in the previous section, \eqref{eq:opt}, was implemented in MATLAB. In order to solve the semidefinite convex optimization problem, the YALMIP framework \cite{lofberg2004yalmip} was used. In all the experiments, the optimization is initialized multiple times, and the best resulting run is used for performance analysis. The source code for learning (MATLAB) and execution (C++) of the Augmented Joint-Space Task-Oriented Dynamical System together with data generated in these experiments are available on-line: \\
\footnotesize Learning:  \underline{\url{https://github.com/epfl-lasa/JT-DS-Learning}}\\
\footnotesize Execution:  \underline{\url{https://github.com/epfl-lasa/JT-DS-lib}}\\
\normalsize

\begin{table}[t]
\centering
\scriptsize
\caption{Datasets used to Evaluate JT-DS learning approach }
\label{tab:datasets}
\begin{tabular}{l|l|l}
\hline\hline
Task Behavior Dataset            & \# Demos & \# Samples (25Hz) \\ \hline
Backhand Stroke     & 11       & 1223                    \\
Forehand Stroke     & 10       & 1424                    \\
Pouring - Free      & 9        & 1032                    \\
Pouring - Obstacle  & 11       & 1232                    \\
Foot Swing          & 8        & 1058                    \\
Singularity Motions & 10       & 1467                 \\ \hline\hline
\end{tabular}
\end{table}

To evaluate the proposed JT-DS learning algorithm we collected demonstrations from 6 different tasks (see Table \ref{tab:datasets}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space. These $M$ demonstrations consist of ($14$-D) time-series $\boldsymbol{\xi}^{(i)} = \{\xi_1^{(i)},\dots,\xi_T^{(i)}\}$ for $i= \{1,\dots,M\}$. The $14$ dimensions in $\xi_t = [q_t,\dot{q_t}]^T$ correspond to joints-space positions and velocities.

\textcolor{red}{2. Table with Evaluation of Learning Schemes for JTDS Models on Different Datasets and discussion of results}
\newpage
\subsection{JT-DS Performance Evaluation}
The performance of the proposed framework was evaluated on a 7-DOF robot arm, the KUKA LWR 4+. The robot is controlled on the joint position level (linearly interpolating from joint velocities) at a rate of $ 500 $ Hz. The resultant joint angles are filtered by a critically damped filter to avoid high torques. Our empirical validation consists of three sections, each highlighting a different advantage of the proposed dynamical system: (i) moving in singular configurations, (ii) following desired behaviors in joint and task spaces simultaneously, and (iii) fast computation/convergence time.

@@

Finally, we are currently working on improving the performance of this method \eqref{eq:ds} by considering a the second order dynamical system. This would provide numerous improvements: a guaranteed smoothness in the velocity profile, the ability to integrate critically damped filter, and the versatility to learn more complex behaviors. \textcolor{red}{Are we actually still doing this?}
%\footnotesize
\appendices

>>>>

\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{url}
\usepackage[font={footnotesize}]{caption}
\usepackage[font={footnotesize}]{subcaption}
\usepackage[usenames, dvipsnames]{color}
\usepackage{amsmath}
\renewcommand{\labelenumii}{\Roman{enumii}} 
\usepackage{authblk}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage[hidelinks]{hyperref}
% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
\usepackage{amssymb}

\usepackage{xcolor,colortbl}

\newcommand{\mc}[2]{\multicolumn{#1}{c}{#2}}
\definecolor{Gray}{gray}{0.85}
\definecolor{LightCyan}{rgb}{0.88,1,1}

@@

%Lastly, we wish to capture joint-space behaviors in our dynamical system.
We make two decisions by the choice of behavior error metric \eqref{eq:optim_error}. First, the metric implicitly suggests that behaviors are defined by expert motions of the behavior, and that fulfilling a behavior means moving similarly to the experts' motion. Second, the type of feature of the trajectory that is being minimized is important. If instead the error feature were joint position distance $\sum_i\left \| q_{d,i} - q_i \right \|$ where $q_d$ is a vector of demonstrated positions, then executing a "behavior" would imply mimicking position, but not velocity. Alternatively, if the error feature were joint velocity direction $\sum_i \left | \frac{\dot{q}_{d,i}}{\left \| \dot{q}_{d,i} \right \|} - \frac{\dot{q}_i}{\left \| \dot{q}_i \right \|} \right |$ mimicking a "behavior" would involve following the motion profile, but not the motion's speed (so for example a slap and a push would exhibit the same behavior). By choosing the combined direction and magnitude of the joint velocity as the error, we are choosing to mimic the magnitude and direction of the motion, which \cite{gielniak2010stylized} suggests is visually most similar to the human definition of "joint motion style".
\section{Augmented \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem} \label{Sec:DS}
\label{sec:proposed_system}
\textcolor{red}{Change semi-definite to definite everywhere\\Change overall $A(q)$ to $\mathcal{A}(q)$\\ Introduce 'behavior synergies' here\\}
We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the two criteria presented in \ref{eq:Lyp} and \ref{eq:optim_error}.
\begin{equation}
\label{eq:ds}
\dot{q} = f(q) = -A(q)J^T(q)P(H(q) - x^*)
\end{equation}
where $P\in \mathbb{R}^{d\times d}$, and $A(q)\in \mathbb{R}^{m\times m}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $A(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy. 
\begin{equation}
\label{eq:A_def}
A(q) = \sum_{k=1}^{K}\theta_k(q)A_k 
\end{equation}
where $A_k\in \mathbb{R}^{m\times m} $ are the different synergies and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are the scheduling parameters\footnote{The scheduling parameters can be a function of time $t$, the states of the
system $q $ or external signals $d(t)$, i.e. $\theta_k (t, q(t), d(t))$. In this paper, we only consider it as a function of the states of the system. It is noteworthy that the presented stability proof can easily be extended for $\theta_k (t, q(t), d(t))$.} modulating each local synergy through time as well as space. Based on \eqref{eq:ds}, one can utilize different synergies in different regions, and compose them to create a more complex multi-behavior motion.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{Pic/A_comparison_positions.pdf}
\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $A$ (emphasizing the hip, knee, and ankle respectively), A: $A(q)=diag(5,1,1) $, B: $A(q)=diag(1,5,1)$, and C: $A(q)=diag(1,1,5)$. On the left, the task-space traces of each motion. On the right, the time-scaled joint positions of each joint. Each motion tends to use its ``primary" joint most and uses the other available joints to compensate for what the primary joint cannot do.}
\label{fig:A_example}
\vspace{-15pt}
\end{figure}

Before proving that the proposed dynamical system satisfies all of the criteria, let us establish an intuitive understanding of the components of the system. While it may seem daunting at first, each of the elements has a straightforward explanation. One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t. the task target (warped according to P), and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}). The positive semi-definite matrix $A(q)$ warps the resulting joint-space velocity; Fig. \ref{fig:A_example} illustrates the effects of  $A(q)$ on the generated motion. Thus the controller can be thought of as a proportional controller in joint space. Lastly, we refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(because it augments the task error, and thus the direction of motion) and $A(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (because it augments the outputted joint velocities).
%When using each of these matrices, the controller will augment the joint velocities to favor using one joint more than the others (in $A_1$ the hip, in $A_2$ the knee, and in $A_3$ the ankle). This ''augmentation" is simply a matrix multiplication ($A\dot{q}_{raw}$). The components of the outputted velocity are linear combinations of the ''raw" un-augmented joint velocities, where the entries in $A$ define the precise ratios. We can see the joint positions resulting from our experiment over time on the right side of Fig. \ref{fig:A_example}. Each motion tends to use its ``primary" joint most (the joint where the velocity is heavily scaled up) and uses the other available joints to compensate for what the primary joint cannot do. For example, Motion 1 starts out using mostly the hip, but once moving the hip no longer brings the system closer to the goal, the system uses the other joints to fold inwards and reach the goal. Ultimately, all three controllers converge on their target, but each has a different path through joint space depending on which joints we choose to emphasize in the behavior matrix $A$. The off-diagonal terms of $A$ have a more nuanced contribution; e.g. $A_{1,2}$ defines the amount that elbow velocity should be converted into hip velocity.

@@

%By learning a position-dependent joint augmentation matrix $A(q)$, we are able to execute the joint behaviors discussed earlier by augment the relative velocities of our joints. \eqref{eq:A_def} constructs our overall $A(q)$ as a linear combination of $A_k~\forall k\in\{1,\dots,K\}$ (each weighted by $\theta_k(q)$), where one can enforce different local joint behaviors in different regions, and compose them to create a more complex multi-behavior motion. In this paper the joint behavior matrix $A(q)$ is determined using only the joint position $q$, though one could potentially use other information such as joint velocity $\dot{q}$, environment force or target task position $x^*$.\\

\section{Learning \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem} 
\label{Sec:Learning}

The behavior of the JT-DS algorithm can be best understood through the lens of synergy control \cite{petrivc2013muscle}. In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the actuation-space that are sufficient to accurately recreate the desired robotic behaviors. In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and $A(q)$ represents the resulting motion constructed from a superposition of different synergies (through \eqref{eq:ds}).

@@

\subsection{Estimating the Synergy Matrices}
Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $A(q)$ as a linear combination of local $A_k$ matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows: 
\begin{equation}
\label{eq:A}
A(q) = \frac{\sum_{k=1}^K A_k \pi_k\mathcal{N}(\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}.
\end{equation} 
Notice the resemblance of \eqref{eq:A} to the Nadaraya-Watson kernel estimator \cite{nadaraya1964regress,watson1964regress}\footnote{The Nadaraya-Watson kernel estimator is used to estimate an unknown regressive function $m(x) = \mathbb{E}\{Y|X\}$, which takes the general form of $\widehat{m}(x) = \frac{\sum_{i=1}^n y_i \mathcal{K}(x,x_i)}{\sum_{i=1}^n \mathcal{K}(x,x_i)}$ where $\mathcal{K}(x,x_i)$ is a kernel function denoting the distance or similarity of $x_i$ to the given location $x$. \cite{nadaraya1964regress,watson1964regress}} with a Gaussian pdf as its kernel function. Hence \eqref{eq:A} can be considered a type of kernel estimator, with the key distinction that the weighting functions $\theta_k'(\phi(q))$ are not determined by individual points (as in the original Nadaraya-Watson kernel estimator) but by the components of a GMM, similar to the weighting functions derived in Gaussian Mixture Regression (GMR) \cite{sung2004gmr}.

@@

\clearpage
\section{Experimental Validation} \label{Sec:Exp}
\textcolor{red}{We should try to have page 5 start with this section, i.e. we must reduce 1 full column from the previous sections.}
\subsection{JT-DS Learning Performance  Evaluation}
To evaluate the proposed JT-DS learning algorithm we collected kinesthetic demonstrations on a 7-DOF KUKA LWR 4+ robot arm, for 6 different tasks (see Table \ref{tab:learning}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space. These behaviors include: a (1) backhand and (2) forehand stroke, a pouring motion (3) with and (4) without an obstacle in its workspace, (5) a foot swing and (6) singularity motions. For each behavior, we collected $M$ demonstrations consisting of ($14$-D) time-series $\boldsymbol{\xi}^{(i)} = \{\xi_1^{(i)},\dots,\xi_T^{(i)}\}$ for $i= \{1,\dots,M\}$. The $14$ dimensions in $\xi_t = [q_t,\dot{q_t}]^T$ correspond to joints angles and velocities. We also record the final task-space configurations ${x^*}^{(i)}$, which we assume is the attractor in task-space. The recordings are initially collected at a rate of 500 Hz. However, our learning approach does not require such a dense representation of the trajectories, hence we down-sample them to 25 Hz. The total number of demonstrations $M$ and samples per \textit{behavior dataset} $M^*T$ are provided in Table \ref{tab:learning}. Given these datasets, we seek to: 
\begin{enumerate}[leftmargin=*]
\item Verify that the learned \textit{behavior synergies} are capable of generating the desired joint-space behaviors.  
\item Find the dimensionality approach which yields the best trade-off between accuracy and model complexity; i.e. the least dimensions $\delta$ and synergies $K$.
\end{enumerate}
We thus evaluate three learning approaches with different dimensionality reduction algorithms: (1) None, (2) PCA and (3) K-PCA with RBF kernel.
We perform 10-fold cross-validation on all datasets, for each learning approach, with a training/testing ratio of $60 \%$. To evaluate the performance of each learning approach, we take the square root of the MSE defined in \eqref{eq:optim_error}, which we refer to as the joint-velocity RMSE. Performing such cross-validation in our case, is however not straight-forward, as our $\mathcal{A}(q)$ formulation has several hyper-parameters, namely the dimensionality $\delta$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$. Moreover, when using K-PCA (with RBF kernel $k(q,q') = \exp(-\frac{||q-q'||^2}{2\sigma^2})$) we must also find the optimal width $\sigma$ in the RBF kernel. Thus, for each fold and each learning approach we find the optimal hyper-parameters $\delta$, $K$ and $\sigma$ (when applicable), as follows:
\textcolor{red}{\begin{enumerate}[leftmargin=*]
\item Explain how optimal $\delta$ is computed for PCA. 
\item Explain how optimal $\delta$ and $\sigma$ are found for K-PCA.
\item Explain how optimal $K$ is computed for GMM.
\end{enumerate}}

\begin{table*}[!t]
\scriptsize
\caption{Performance Comparison of Learning Approach with different Dimensionality Reduction Schemes. We present the \texttt{mean(std)} for the optimal lower-dimensional $d$ and behavior synergies $K$ and RMSE on training and testing set, found for every learning scheme over 10 runs.\label{tab:learning}}
\begin{center}
\begin{tabular}{cc|cc|cc}
    \hline
    \hline
    \multicolumn{1}{c}{\multirow{4}{*}{Behavior Dataset}} & \multicolumn{1}{c}{\multirow{4}{*}{Dim. Red. Approach}} & \multicolumn{4}{c}{\multirow{2}{*}{Optimal Parameters | Joint Velocity RMSE [$rad/s$]}}
	\\  
    & & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $\delta$}} & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $K$}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Train}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Test}}	\\ \\ \hline

    \multicolumn{1}{c}{\multirow{3}{*}{Backhand Stroke}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
    \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
	& K-PCA ($\sigma = 6.238$) & \textbf{3.1 $\pm$ (0.316)} & \textbf{3.1 $\pm$ (0.316)} & 0.219 $\pm$ (0.013) & \cellcolor{blue!15} 0.355 $\pm$ (0.068)   \\
	\hline

	\multicolumn{1}{c}{\multirow{3}{*}{Forehand Stroke}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
		& K-PCA ($\sigma = 5.488$) & \textbf{3.9 $\pm$ (0.316)} & \textbf{3.2 $\pm$ (0.421)} & 0.305 $\pm$ (0.021) & \cellcolor{blue!15} 0.433 $\pm$ (0.104)  \\
		\hline

	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Free}} & None & 7 & \textbf{ 4.1 $\pm$ (1.286)} & 0.163 $\pm$ (0.026) & 7.609 $\pm$ (18.99) \\
   \multicolumn{1}{c}{\multirow{3}{*}{($M=9,M^*T=1032$)}} & PCA & \textbf{ 2.3 $\pm$ (0.483)} & 4.8 $\pm$ (2.044) & \textbf{ 0.161 $\pm$ (0.034) } & \cellcolor{blue!10} 0.623 $\pm$ (0.301) \\	
	& K-PCA ($\sigma = 3.92$) & 3 & 4.3 $\pm$ (1.767) & 0.171 $\pm$ (0.106) & \cellcolor{blue!15} 0.511 $\pm$ (0.305)   \\
				\hline
	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Obstacle}} & None & 7 & 6.2 $\pm$ (2.573) & \textbf{0.247 $\pm$ (0.039)} & 3.049 $\pm$ (3.113) \\
   \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1232$)}} & PCA & 3.8 $\pm$ (0.421) & 3.8 $\pm$ (1.549) & 0.292 $\pm$ (0.018) & \cellcolor{blue!10} 0.486 $\pm$ (0.125) \\	
				& K-PCA ($\sigma = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & 3.1 $\pm$ (0.316) & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
				\hline	
	\multicolumn{1}{c}{\multirow{3}{*}{Foot Swing}} & None & 7 & 4.4 $\pm$ (2.366) & \textbf{ 0.080 $\pm$ (0.008) } & 0.368 $\pm$ (0.265)  \\
	\multicolumn{1}{c}{\multirow{3}{*}{($M=8,M^*T=1058$)}} & PCA & \textbf{1} & 3.2 $\pm$ (0.421) & 0.084 $\pm$ (0.004) & \cellcolor{blue!10}  0.123 $\pm$ (0.011) \\	
	& K-PCA ($\sigma = 1.513$) & 3.2 $\pm$ (0.421) & \textbf{3} &  0.091 $\pm$ (0.007) & \cellcolor{blue!15} 0.115 $\pm$ (0.015)   \\
				\hline

	\multicolumn{1}{c}{\multirow{3}{*}{Singularity Motions}} & None & 7 & 6.2 $\pm$ (2.898) & 0.054 $\pm$ (0.009) & 0.137 $\pm$ (0.098)  \\
    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1467$)}} & PCA & \textbf{2} & 6 $\pm$ (2.357) & \textbf{ 0.053 $\pm$ (0.010) } & \cellcolor{blue!10} 0.105 $\pm$ (0.046) \\	
	& K-PCA ($\sigma = 1.769$) & 3.6 $\pm$ (0.516) & \textbf{5.3 $\pm$ (2.452)} & 0.055 $\pm$ (0.018) & \cellcolor{blue!15} 0.071 $\pm$ (0.010)   \\ \hline\hline
\end{tabular}
\end{center}
\vspace{-10pt}

\end{table*}

Once these optimal hyper-parameters are estimated for each fold, we solve for the definite convex optimization problem in order to find our synergy matrices $A$'s which best minimize the objective function \eqref{eq:opt}. We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from the training/testing datasets, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing set trajectories.

\textcolor{red}{Discuss results: As can be seen in Table \ref{tab:learning}, ...}

This learning pipeline was implemented in MATLAB. In order to solve the definite convex optimization problem, the YALMIP framework \cite{lofberg2004yalmip} was used. In all the experiments, the optimization is initialized multiple times, and the best resulting run is used for performance analysis. The source code for learning (MATLAB) and execution (C++) of the Augmented Joint-Space Task-Oriented Dynamical System together with data generated in these experiments are available on-line: \\
\footnotesize Learning:  \underline{\url{https://github.com/epfl-lasa/JT-DS-Learning}}\\
\normalsize
\textcolor{red}{In bla, we provide the necessary code to run this learning approach on our test datasets as well as on your own. We also provide code for the cross-validation evaluation.}

\subsection{Comparison with Cartesian DS + IK Solver}
We now evaluate the performance of the properties of JT-DS\footnote{With the optimal learning approach found in the previous section.} against a standard Cartesian motion generator + IK solver approach. In this evaluation we seek to compare the capabilities (or lack thereof) of both approaches to: 
\begin{enumerate}
\item Follow desired behaviors in both joint and task spaces.
\item Achieve fast computation/convergence time.
\item Transit through singular configurations. 
\end{enumerate}
We continue our evaluation on a 7-DOF robot arm, the KUKA LWR 4+. The robot is controlled on the joint position level (linearly interpolating from joint velocities) at a rate of $ 500 $ Hz. The resultant joint angles are filtered by a critically damped filter to avoid high torques.

\begin{figure}[t]
\vspace{-5pt}
	\centering
	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
	\caption{End-effector trajectories for the footstep motion in Cartesian space. The JT-DS motion moves smoothly closely resembles the demonstrated trajectories. On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing. \label{fig:foot}}
\end{figure}

\subsubsection{Following Desired Joint Behaviors}
To evaluate the system's ability to track learned joint behaviors, we devised two experiments (summarized in Table \ref{table:1}) comparing the tracking capabilities of our JT-DS method with those of the Cartesian-space SEDS \cite{khansari2011learning} algorithm using damped least square IK. In the first experiment, the robot was guided through a complex motion task, moving an object through an environment with an obstacle. 
Human experts guided the robot through a series of demonstrations, ensuring that neither the robot's joints nor the held object intersected the obstacle. The results can be seen in Fig. \ref{fig:snapshot:a} and Fig. \ref{fig:snapshot:b}. The JT-DS algorithm mimicked the joint-space behaviors of the demonstration (e.g. folding the elbow, lowering the shoulder), and therefore successfully avoided the obstacle while still converging to the desired Cartesian position. Meanwhile, SEDS only learned the demonstrated behavior in task-space (the position of the end-effector is signified by a black ball) meaning that the robot was not constrained in joint space. This ultimately led to one of its joints colliding with the obstacle. It should be noted that the JT-DS motion did not follow the demonstrations in task-space very closely (as we would expect), but did ultimately converge to its target position.
In the second experiment the robot was taught a footstep-like motion, beginning with a straight leg, moving through a singularity, and finally bringing the knee up (Fig. \ref{fig:snapshot:c}). JT-DS followed the demonstrations closely, while SEDS became unstable in the singularity (Fig. \ref{fig:foot}). This demonstrates the JT-DS algorithm's ability to move cleanly in and out of singularities, overcoming the Achilles heel of equivalent Cartesian-based motion generators.

\begin{table}[t]
	\tiny
	\vspace{0.1cm}
	\caption{  Each simulated trajectory is initialized at $q=[0~\dots~0]^T$. The convergence duration is the time required to move within $0.001$m of the target. The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.}
	\label{table:2}
	\centering
	\begin{tabular}{l|l|l|l}
		\hline\hline
		& Norm. converg. [{s}/{m}] & Comp. time {[}ms{]} & Track. error {[}m{]} \\ \hline
		SEDS  + IK    & $10.6566\pm2.1364$                          & $71.7752\pm4.9189$       & $0.0163\pm0.0063$      \\ 
		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline\hline
	\end{tabular}
	\hspace{0.1cm}
\end{table}

@@

%Finally, we are currently working on improving the performance of this method \eqref{eq:ds} by considering a the second order dynamical system. This would provide numerous improvements: a guaranteed smoothness in the velocity profile, the ability to integrate critically damped filter, and the versatility to learn more complex behaviors. \textcolor{red}{Are we actually still doing this?}
%\footnotesize
\appendices
